### 2017-11-06 submission/manuscript.Rmd
add total numbers to background

Using both fecal and tissue samples totalling over 2100 total individuals across 14 studies [@hale_shifts_2017; @flemer_tumour-associated_2017; @chen_human_2012; @wang_structural_2012; @lu_mucosal_2016; @brim_microbiome_2013; @weir_stool_2013; @dejea_microbiota_2014; @sanapareddy_increased_2012; @ahn_human_2013; @burns_virulence_2015; @zeller_potential_2014; @baxter_microbiota-based_2016; @geng_diversified_2013] within our data analysis we expand both the breadth and scope of the previous meta-analysis to investigate whether the bacterial community is an important risk factor for both adenoma and carcinoma. We first assessed the diversity of controls, adenoma, and carcinoma individuals and tested whether they change and if it results in an increased relative risk of adenoma or carcinoma. Next, we assessed how this relative risk compared to CRC associated genera for both adenoma and carcinoma. Third, using Random Forest models we assessed whether the community context can increase the classification model area under the curve (AUC). Finally, we examine whether the studies that were used were adequately powered and if not what effect size they were powered for.

>>>>

Using both fecal (total individuals = `r sum(stool_n$control) + sum(stool_n$adenoma, na.rm = T) + sum(stool_n$cancer, na.rm = T)`) and tissue samples (total samples = `r sum(tissue_n$control) + sum(tissue_n$adenoma, na.rm = T) + sum(tissue_n$cancer, na.rm = T)`) totalling over `r sum(stool_n$control) + sum(stool_n$adenoma, na.rm = T) + sum(stool_n$cancer, na.rm = T) + sum(tissue_n$control) + sum(tissue_n$adenoma, na.rm = T) + sum(tissue_n$cancer, na.rm = T)` total samples across 14 studies [@hale_shifts_2017; @flemer_tumour-associated_2017; @chen_human_2012; @wang_structural_2012; @lu_mucosal_2016; @brim_microbiome_2013; @weir_stool_2013; @dejea_microbiota_2014; @sanapareddy_increased_2012; @ahn_human_2013; @burns_virulence_2015; @zeller_potential_2014; @baxter_microbiota-based_2016; @geng_diversified_2013] within our data analysis we expand both the breadth and scope of the previous meta-analysis to investigate whether the bacterial community is an important risk factor for both adenoma and carcinoma. We first assessed the diversity of controls, adenoma, and carcinoma individuals and tested whether they change and if it results in an increased relative risk of adenoma or carcinoma. Next, we assessed how this relative risk compared to CRC associated genera for both adenoma and carcinoma. Third, using Random Forest models we assessed whether the community context can increase the classification model area under the curve (AUC). Finally, we examine whether the studies that were used were adequately powered and if not what effect size they were powered for.

### 2017-11-06 submission/manuscript.tex
add total numbers to background

Using both fecal and tissue samples totalling over 2100 total
individuals across 14 studies {[}12--25{]} within our data analysis we
expand both the breadth and scope of the previous meta-analysis to
investigate whether the bacterial community is an important risk factor
for both adenoma and carcinoma. We first assessed the diversity of
controls, adenoma, and carcinoma individuals and tested whether they
change and if it results in an increased relative risk of adenoma or
carcinoma. Next, we assessed how this relative risk compared to CRC
associated genera for both adenoma and carcinoma. Third, using Random
Forest models we assessed whether the community context can increase the
classification model area under the curve (AUC). Finally, we examine
whether the studies that were used were adequately powered and if not
what effect size they were powered for.

>>>>

Using both fecal (total individuals = 1737) and tissue samples (total
samples = 492) totalling over 2229 total samples across 14 studies
{[}12--25{]} within our data analysis we expand both the breadth and
scope of the previous meta-analysis to investigate whether the bacterial
community is an important risk factor for both adenoma and carcinoma. We
first assessed the diversity of controls, adenoma, and carcinoma
individuals and tested whether they change and if it results in an
increased relative risk of adenoma or carcinoma. Next, we assessed how
this relative risk compared to CRC associated genera for both adenoma
and carcinoma. Third, using Random Forest models we assessed whether the
community context can increase the classification model area under the
curve (AUC). Finally, we examine whether the studies that were used were
adequately powered and if not what effect size they were powered for.
